% !TeX TS-program = lualatex
% !TeX encoding = UTF-8
\documentclass[../interior-demo.tex]{subfiles}
\begin{document}

\begin{opening}
\null\null\null\null\null
\name{3. The Mississippi River}
\null
\subname{Excerpt from \ital{Life on the Mississippi}, 1883}
\null\null\null\null\null
\end{opening}

\init[-.1]{T}{he Mississippi} is well worth reading about. It is not a
commonplace river, but on the contrary is in all ways remarkable.

\scene{+I. As a River}

Considering the
Missouri its main branch, it is the longest river in the world---four
thousand three hundred miles. It seems safe to say that it is also the
crookedest river in the world, since in one part of its journey it uses
up one thousand three hundred miles to cover the same ground that the
crow would fly over in six hundred and seventy-five.

It discharges three
times as much water as the St. Lawrence, twenty-five times as much
as the Rhine, and three hundred and thirty-eight times as much as the
Thames. No other river has so vast a drainage-basin. The
Mississippi receives and carries to the Gulf water from fifty-four
subordinate rivers that are navigable by steamboats, and from some
hundreds that are navigable by flats and keels. The area of its
drainage-basin is as great as the combined areas of England, Wales,
Scotland, Ireland, France, Spain, Portugal, Germany, Austria, Italy,
and Turkey; and almost all this wide region is fertile.

It is a remarkable river in this: that instead of widening toward its
mouth, it grows narrower; grows narrower and deeper. From the junction
of the Ohio to a point half way down to the sea, the width averages a
mile in high water: thence to the sea the width steadily diminishes,
until, at the 'Passes,' above the mouth, it is but little over half
a mile. At the junction of the Ohio the Mississippi's depth is
eighty-seven feet; the depth increases gradually, reaching one hundred
and twenty-nine just above the mouth.

The difference in rise and fall is also remarkable---not in the upper,
but in the lower river. The rise is tolerably uniform down to Natchez
(three hundred and sixty miles above the mouth)---about fifty feet.
But at Bayou La Fourche the river rises only twenty-four feet; at New
Orleans only fifteen, and just above the mouth only two and one half.

An article in the New Orleans 'Times-Democrat,' based upon reports of
able engineers, states that the river annually empties four hundred and
six million tons of mud into the Gulf of Mexico---which brings to mind
Captain Marryat's rude name for the Mississippi---'the Great Sewer.' This
mud, solidified, would make a mass a mile square and two hundred and
forty-one feet high.

The mud deposit gradually extends the land---but only gradually; it has
extended it not quite a third of a mile in the two hundred years which
have elapsed since the river took its place in history. The belief of
the scientific people is, that the mouth used to be at Baton Rouge,
where the hills cease, and that the two hundred miles of land between
there and the Gulf was built by the river. This gives us the age of that
piece of country, without any trouble at all---one hundred and twenty
thousand years.

The Mississippi is remarkable in still another way---its disposition to
make prodigious jumps by cutting through narrow necks of land, and thus
straightening and shortening itself. More than once it has shortened
itself thirty miles at a single jump! These cut-offs have had curious
effects: they have thrown several river towns out into the rural
districts, and built up sand bars and forests in front of them. The town
of Delta used to be three miles below Vicksburg: a recent cutoff has
radically changed the position, and Delta is now \ital{two miles above}
Vicksburg. Both of these river towns have been retired to the country by that
cut-off.

The Mississippi does not alter its locality by cut-offs alone: it
is always changing its habitat \ital{bodily}---is always moving bodily
\ital{sidewise}. At Hard Times, La., the river is two miles west of the
region it used to occupy. As a result, the original site of that
settlement is not now in Louisiana at all, but on the other side of
the river, in the State of Mississippi. \ital{Nearly the whole of that one
thousand three hundred miles of old Mississippi River which La Salle
floated down in his canoes, two hundred years ago, is good solid dry
ground now}. The river lies to the right of it, in places, and to the
left of it in other places.

Although the Mississippi's mud builds land but slowly, down at the
mouth, where the Gulfs billows interfere with its work, it builds fast
enough in better protected regions higher up: for instance, Prophet's
Island contained one thousand five hundred acres of land thirty years
ago; since then the river has added seven hundred acres to it.

\scene{+II. Early History}

Let us drop the Mississippi's physical history, and say a word about its
historical history---so to speak.

The world and the books are so accustomed to use, and over-use, the word
'new' in connection with our country, that we early get and permanently
retain the impression that there is nothing old about it. We do of
course know that there are several comparatively old dates in American
history, but the mere figures convey to our minds no just idea, no
distinct realization, of the stretch of time which they represent.
To say that De Soto, the first European who ever saw the Mississippi
River, saw it in 1542, is a remark which states a fact without
interpreting it.

The date 1542, standing by itself, means little or nothing to us; but
when one groups a few neighboring historical dates and facts around it,
he adds perspective and color, and then realizes that this is one of the
American dates which is quite respectable for age.

For instance, when the Mississippi was first seen by a European, less
than a quarter of a century had elapsed since the death of Raphael;
the driving out of the Knights-Hospitallers from Rhodes by
the Turks; and the placarding of the Ninety-Five Propositions.
When De Soto took his glimpse of the river,
Michael Angelo's paint was not yet dry on the Last
Judgment in the Sistine Chapel; Elizabeth of England was not yet in her teens;
'Don Quixote' was not yet written; Shakespeare was not yet born;
a hundred long years must
still elapse before Englishmen would hear the name of Oliver Cromwell.

Unquestionably the discovery of the Mississippi is a datable fact which
considerably mellows and modifies the shiny newness of our country, and
gives her a most respectable outside-aspect of rustiness and antiquity.

De Soto merely glimpsed the river, then died and was buried in it by his
priests and soldiers. One would expect the priests and the soldiers
to multiply the river's dimensions by ten---the Spanish custom of the
day---and thus move other adventurers to go at once and explore it. On
the contrary, their narratives when they reached home, did not excite
that amount of curiosity. The Mississippi was left unvisited by Europeans
during a term of years which seems incredible in our energetic days.

One may 'sense' the interval to his mind, after a fashion, by dividing it
up in this way: After De Soto glimpsed the river, a fraction short of
a quarter of a century elapsed, and then Shakespeare was born; lived a
trifle more than half a century, then died; and when he had been in his
grave considerably more than half a century, the \ital{second} European saw
the Mississippi. In our day we don't allow a hundred and thirty years to
elapse between glimpses of a marvel. If somebody should discover a creek
in the county next to the one that the North Pole is in, Europe and
America would start fifteen costly expeditions thither: one to explore
the creek, and the other fourteen to hunt for each other.


\end{document}
